---
title: "Fit SIRS to RV cases (simulated data #2) -- sensitivity analysis"
output: pdf_document
date: "2026-07-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir=".../Data_and_Codes")
```

```{r}
rm(list=ls())

library(tidyverse)
library(dplyr)
library(cowplot)
library(rstan)
```

```{r}
n_chains <- 4
n_warmup <- 2000
n_iter <- n_warmup + 2000

gamma_rv <- log(1/(1-7/8)) # approx 8 days on average
mu <- 1/80/52
pop <- 5E+6
```

```{r}
plot_fit <- function(posterior, data){
  return(
    data.frame("time"=1:data$n_week,
               "observed"=data$obs,
               "fitted"=apply(posterior$cases, 2, median),
               "lower"=apply(posterior$cases, 2, quantile, probs=0.025),
               "upper"=apply(posterior$cases, 2, quantile, probs=0.975)) %>%
      ggplot(aes(x=time)) +
      geom_ribbon(aes(ymin=lower, ymax=upper), fill="red", alpha=0.5) + # Credible interval
      geom_line(aes(y=fitted), color = "red", linewidth=0.1) + # Posterior median
      geom_point(aes(y = observed), cex=0.5) + # Observed cases
      theme_minimal() +
      theme(axis.title.x=element_blank())
  )
}

plot_beta <- function(posterior){
  return(
    posterior$beta %>% apply(2, function(X){
      return(data.frame("mean"=mean(X),
                        "CI95_lower"=quantile(X, probs=0.025),
                        "CI95_upper"=quantile(X, probs=0.975), row.names=NULL))
  }) %>% bind_rows %>% cbind("week"=1:52,.) %>%
    ggplot(aes(x=week, y=mean)) +
    geom_ribbon(aes(ymin=CI95_lower, ymax=CI95_upper), fill="red", col=NA, alpha=0.5) +
    geom_line(aes(y=mean), col="red") +
    theme_classic()
  )
}
```

```{r}
parms <- c("logbeta0", "eps_beta[1]", "eps_beta[2]", "eps_beta[3]", "eps_beta[4]", "sigma_beta",
           "kappa", "phi", "omega", "rho", "S0", "I0", "sigma")

model <- stan_model("Fit/Stan_models/SIRS_npi_vi.stan")
```

# Omega[1]

## Import and plot simulated data

```{r}
load("Data/Simulation/cases_2_omega1.RData")
real_parms <- readRDS("Data/Simulation/parms_simul_2_omega1.rds")

plot_grid(
  ggplot(obs_rv, aes(x=time/52)) +
    geom_point(aes(y=cases_noisy)) +
    geom_line(aes(y=cases)) +
    scale_x_continuous(breaks=0:10),
  ggplot(obs_iav, aes(x=time/52)) +
    geom_point(aes(y=cases_noisy)) +
    geom_point(aes(y=smooth_cases_noisy), col="red") +
    geom_line(aes(y=cases)) +
    scale_x_continuous(breaks=0:10),
  ncol=1)
```

## Fit all data with potential viral interaction

```{r}
data_stan <- list(
  n_week=nrow(obs_rv),
  week=obs_rv$week,
  obs=obs_rv$cases_noisy,
  obs_IAV=obs_iav$smooth_cases_noisy/max(obs_iav$smooth_cases_noisy),
  c=obs_rv$c,
  gamma=gamma_rv,
  mu=mu,
  pop=pop
)
```

```{r, eval=FALSE}
fit <- sampling(model, data=data_stan,
                chains=n_chains, iter=n_iter, warmup=n_warmup, seed=1001,
                control=list(adapt_delta=0.95, max_treedepth=12),
                cores=n_chains)

saveRDS(fit, "Fit/Simulation_study/Output/fit_RV_SIRS_npi_vi_simul2_omega1.RDS")
```

```{r}
fit <- readRDS("Fit/Simulation_study/Output/fit_RV_SIRS_npi_vi_simul2_omega1.RDS")

check_hmc_diagnostics(fit)
get_num_divergent(fit)
get_num_max_treedepth(fit)
get_low_bfmi_chains(fit)
get_bfmi(fit)

summary(fit)$summary[c(1:5,53:60),]
```

```{r}
posterior <- extract(fit)

traceplot(fit, pars=parms)
pairs(fit, pars=parms)
plot_fit(posterior, data_stan)
plot_beta(posterior)
```

# Omega[2]

## Import and plot simulated data

```{r}
load("Data/Simulation/cases_2_omega2.RData")
real_parms <- readRDS("Data/Simulation/parms_simul_2_omega2.rds")

plot_grid(
  ggplot(obs_rv, aes(x=time/52)) +
    geom_point(aes(y=cases_noisy)) +
    geom_line(aes(y=cases)) +
    scale_x_continuous(breaks=0:10),
  ggplot(obs_iav, aes(x=time/52)) +
    geom_point(aes(y=cases_noisy)) +
    geom_point(aes(y=smooth_cases_noisy), col="red") +
    geom_line(aes(y=cases)) +
    scale_x_continuous(breaks=0:10),
  ncol=1)
```

## Fit all data with potential viral interaction

```{r}
data_stan <- list(
  n_week=nrow(obs_rv),
  week=obs_rv$week,
  obs=obs_rv$cases_noisy,
  obs_IAV=obs_iav$smooth_cases_noisy/max(obs_iav$smooth_cases_noisy),
  c=obs_rv$c,
  gamma=gamma_rv,
  mu=mu,
  pop=pop
)
```

```{r, eval=FALSE}
fit <- sampling(model, data=data_stan,
                chains=n_chains, iter=n_iter, warmup=n_warmup, seed=1001,
                control=list(adapt_delta=0.95, max_treedepth=12),
                cores=n_chains)

saveRDS(fit, "Fit/Simulation_study/Output/fit_RV_SIRS_npi_vi_simul2_omega2.RDS")
```

```{r}
fit <- readRDS("Fit/Simulation_study/Output/fit_RV_SIRS_npi_vi_simul2_omega2.RDS")

check_hmc_diagnostics(fit)
get_num_divergent(fit)
get_num_max_treedepth(fit)
get_low_bfmi_chains(fit)
get_bfmi(fit)

summary(fit)$summary[c(1:5,53:60),]
```

```{r}
posterior <- extract(fit)

traceplot(fit, pars=parms)
pairs(fit, pars=parms)
plot_fit(posterior, data_stan)
plot_beta(posterior)
```

# kappa_rv[1]

## Import and plot simulated data

```{r}
load("Data/Simulation/cases_2_kappa_rv1.RData")
real_parms <- readRDS("Data/Simulation/parms_simul_2_kappa_rv1.rds")

plot_grid(
  ggplot(obs_rv, aes(x=time/52)) +
    geom_point(aes(y=cases_noisy)) +
    geom_line(aes(y=cases)) +
    scale_x_continuous(breaks=0:10),
  ggplot(obs_iav, aes(x=time/52)) +
    geom_point(aes(y=cases_noisy)) +
    geom_point(aes(y=smooth_cases_noisy), col="red") +
    geom_line(aes(y=cases)) +
    scale_x_continuous(breaks=0:10),
  ncol=1)
```

## Fit all data with potential viral interaction

```{r}
data_stan <- list(
  n_week=nrow(obs_rv),
  week=obs_rv$week,
  obs=obs_rv$cases_noisy,
  obs_IAV=obs_iav$smooth_cases_noisy/max(obs_iav$smooth_cases_noisy),
  c=obs_rv$c,
  gamma=gamma_rv,
  mu=mu,
  pop=pop
)
```

```{r, eval=FALSE}
fit <- sampling(model, data=data_stan,
                chains=n_chains, iter=n_iter, warmup=n_warmup, seed=1001,
                control=list(adapt_delta=0.95, max_treedepth=12),
                cores=n_chains)

saveRDS(fit, "Fit/Simulation_study/Output/fit_RV_SIRS_npi_vi_simul2_kappa_rv1.RDS")
```

```{r}
fit <- readRDS("Fit/Simulation_study/Output/fit_RV_SIRS_npi_vi_simul2_kappa_rv1.RDS")

check_hmc_diagnostics(fit)
get_num_divergent(fit)
get_num_max_treedepth(fit)
get_low_bfmi_chains(fit)
get_bfmi(fit)

summary(fit)$summary[c(1:5,53:60),]
```

```{r}
posterior <- extract(fit)

traceplot(fit, pars=parms)
pairs(fit, pars=parms)
plot_fit(posterior, data_stan)
plot_beta(posterior)
```

# kappa_rv[2]

## Import and plot simulated data

```{r}
load("Data/Simulation/cases_2_kappa_rv2.RData")
real_parms <- readRDS("Data/Simulation/parms_simul_2_kappa_rv2.rds")

plot_grid(
  ggplot(obs_rv, aes(x=time/52)) +
    geom_point(aes(y=cases_noisy)) +
    geom_line(aes(y=cases)) +
    scale_x_continuous(breaks=0:10),
  ggplot(obs_iav, aes(x=time/52)) +
    geom_point(aes(y=cases_noisy)) +
    geom_point(aes(y=smooth_cases_noisy), col="red") +
    geom_line(aes(y=cases)) +
    scale_x_continuous(breaks=0:10),
  ncol=1)
```

## Fit all data with potential viral interaction

```{r}
data_stan <- list(
  n_week=nrow(obs_rv),
  week=obs_rv$week,
  obs=obs_rv$cases_noisy,
  obs_IAV=obs_iav$smooth_cases_noisy/max(obs_iav$smooth_cases_noisy),
  c=obs_rv$c,
  gamma=gamma_rv,
  mu=mu,
  pop=pop
)
```

```{r, eval=FALSE}
fit <- sampling(model, data=data_stan,
                chains=n_chains, iter=n_iter, warmup=n_warmup, seed=1001,
                control=list(adapt_delta=0.96, max_treedepth=12),
                cores=n_chains)

saveRDS(fit, "Fit/Simulation_study/Output/fit_RV_SIRS_npi_vi_simul2_kappa_rv2.RDS")
```

```{r}
fit <- readRDS("Fit/Simulation_study/Output/fit_RV_SIRS_npi_vi_simul2_kappa_rv2.RDS")

check_hmc_diagnostics(fit)
get_num_divergent(fit)
get_num_max_treedepth(fit)
get_low_bfmi_chains(fit)
get_bfmi(fit)

summary(fit)$summary[c(1:5,53:60),]
```

```{r}
posterior <- extract(fit)

traceplot(fit, pars=parms)
pairs(fit, pars=parms)
plot_fit(posterior, data_stan)
plot_beta(posterior)
```

# kappa_iav[1]

## Import and plot simulated data

```{r}
load("Data/Simulation/cases_2_kappa_iav1.RData")
real_parms <- readRDS("Data/Simulation/parms_simul_2_kappa_iav1.rds")

plot_grid(
  ggplot(obs_rv, aes(x=time/52)) +
    geom_point(aes(y=cases_noisy)) +
    geom_line(aes(y=cases)) +
    scale_x_continuous(breaks=0:10),
  ggplot(obs_iav, aes(x=time/52)) +
    geom_point(aes(y=cases_noisy)) +
    geom_point(aes(y=smooth_cases_noisy), col="red") +
    geom_line(aes(y=cases)) +
    scale_x_continuous(breaks=0:10),
  ncol=1)
```

## Fit all data with potential viral interaction

```{r}
data_stan <- list(
  n_week=nrow(obs_rv),
  week=obs_rv$week,
  obs=obs_rv$cases_noisy,
  obs_IAV=obs_iav$smooth_cases_noisy/max(obs_iav$smooth_cases_noisy),
  c=obs_rv$c,
  gamma=gamma_rv,
  mu=mu,
  pop=pop
)
```

```{r, eval=FALSE}
fit <- sampling(model, data=data_stan,
                chains=n_chains, iter=n_iter, warmup=n_warmup, seed=101,
                control=list(adapt_delta=0.95, max_treedepth=12),
                cores=n_chains)

saveRDS(fit, "Fit/Simulation_study/Output/fit_RV_SIRS_npi_vi_simul2_kappa_iav1.RDS")
```

```{r}
fit <- readRDS("Fit/Simulation_study/Output/fit_RV_SIRS_npi_vi_simul2_kappa_iav1.RDS")

check_hmc_diagnostics(fit)
get_num_divergent(fit)
get_num_max_treedepth(fit)
get_low_bfmi_chains(fit)
get_bfmi(fit)

summary(fit)$summary[c(1:5,53:60),]
```

```{r}
posterior <- extract(fit)

traceplot(fit, pars=parms)
pairs(fit, pars=parms)
plot_fit(posterior, data_stan)
plot_beta(posterior)
```

# kappa_iav[2]

## Import and plot simulated data

```{r}
load("Data/Simulation/cases_2_kappa_iav2.RData")
real_parms <- readRDS("Data/Simulation/parms_simul_2_kappa_iav2.rds")

plot_grid(
  ggplot(obs_rv, aes(x=time/52)) +
    geom_point(aes(y=cases_noisy)) +
    geom_line(aes(y=cases)) +
    scale_x_continuous(breaks=0:10),
  ggplot(obs_iav, aes(x=time/52)) +
    geom_point(aes(y=cases_noisy)) +
    geom_point(aes(y=smooth_cases_noisy), col="red") +
    geom_line(aes(y=cases)) +
    scale_x_continuous(breaks=0:10),
  ncol=1)
```

## Fit all data with potential viral interaction

```{r}
data_stan <- list(
  n_week=nrow(obs_rv),
  week=obs_rv$week,
  obs=obs_rv$cases_noisy,
  obs_IAV=obs_iav$smooth_cases_noisy/max(obs_iav$smooth_cases_noisy),
  c=obs_rv$c,
  gamma=gamma_rv,
  mu=mu,
  pop=pop
)
```

```{r, eval=FALSE}
fit <- sampling(model, data=data_stan,
                chains=n_chains, iter=n_iter, warmup=n_warmup, seed=1001,
                control=list(adapt_delta=0.96, max_treedepth=12),
                cores=n_chains)

saveRDS(fit, "Fit/Simulation_study/Output/fit_RV_SIRS_npi_vi_simul2_kappa_iav2.RDS")
```

```{r}
fit <- readRDS("Fit/Simulation_study/Output/fit_RV_SIRS_npi_vi_simul2_kappa_iav2.RDS")

check_hmc_diagnostics(fit)
get_num_divergent(fit)
get_num_max_treedepth(fit)
get_low_bfmi_chains(fit)
get_bfmi(fit)

summary(fit)$summary[c(1:5,53:60),]
```

```{r}
posterior <- extract(fit)

traceplot(fit, pars=parms)
pairs(fit, pars=parms)
plot_fit(posterior, data_stan)
plot_beta(posterior)
```

# a_iav[1]

## Import and plot simulated data

```{r}
load("Data/Simulation/cases_2_a_iav1.RData")
real_parms <- readRDS("Data/Simulation/parms_simul_2_a_iav1.rds")

plot_grid(
  ggplot(obs_rv, aes(x=time/52)) +
    geom_point(aes(y=cases_noisy)) +
    geom_line(aes(y=cases)) +
    scale_x_continuous(breaks=0:10),
  ggplot(obs_iav, aes(x=time/52)) +
    geom_point(aes(y=cases_noisy)) +
    geom_point(aes(y=smooth_cases_noisy), col="red") +
    geom_line(aes(y=cases)) +
    scale_x_continuous(breaks=0:10),
  ncol=1)
```

## Fit all data with potential viral interaction

```{r}
data_stan <- list(
  n_week=nrow(obs_rv),
  week=obs_rv$week,
  obs=obs_rv$cases_noisy,
  obs_IAV=obs_iav$smooth_cases_noisy/max(obs_iav$smooth_cases_noisy),
  c=obs_rv$c,
  gamma=gamma_rv,
  mu=mu,
  pop=pop
)
```

```{r, eval=FALSE}
fit <- sampling(model, data=data_stan,
                chains=n_chains, iter=n_iter, warmup=n_warmup, seed=1001,
                control=list(adapt_delta=0.95, max_treedepth=12),
                cores=n_chains)

saveRDS(fit, "Fit/Simulation_study/Output/fit_RV_SIRS_npi_vi_simul2_a_iav1.RDS")
```

```{r}
fit <- readRDS("Fit/Simulation_study/Output/fit_RV_SIRS_npi_vi_simul2_a_iav1.RDS")

check_hmc_diagnostics(fit)
get_num_divergent(fit)
get_num_max_treedepth(fit)
get_low_bfmi_chains(fit)
get_bfmi(fit)

summary(fit)$summary[c(1:5,53:60),]
```

```{r}
posterior <- extract(fit)

traceplot(fit, pars=parms)
pairs(fit, pars=parms)
plot_fit(posterior, data_stan)
plot_beta(posterior)
```

# a_iav[2]

## Import and plot simulated data

```{r}
load("Data/Simulation/cases_2_a_iav2.RData")
real_parms <- readRDS("Data/Simulation/parms_simul_2_a_iav2.rds")

plot_grid(
  ggplot(obs_rv, aes(x=time/52)) +
    geom_point(aes(y=cases_noisy)) +
    geom_line(aes(y=cases)) +
    scale_x_continuous(breaks=0:10),
  ggplot(obs_iav, aes(x=time/52)) +
    geom_point(aes(y=cases_noisy)) +
    geom_point(aes(y=smooth_cases_noisy), col="red") +
    geom_line(aes(y=cases)) +
    scale_x_continuous(breaks=0:10),
  ncol=1)
```

## Fit all data with potential viral interaction

```{r}
data_stan <- list(
  n_week=nrow(obs_rv),
  week=obs_rv$week,
  obs=obs_rv$cases_noisy,
  obs_IAV=obs_iav$smooth_cases_noisy/max(obs_iav$smooth_cases_noisy),
  c=obs_rv$c,
  gamma=gamma_rv,
  mu=mu,
  pop=pop
)
```

```{r, eval=FALSE}
fit <- sampling(model, data=data_stan,
                chains=n_chains, iter=n_iter, warmup=n_warmup, seed=1001,
                control=list(adapt_delta=0.96, max_treedepth=12),
                cores=n_chains)

saveRDS(fit, "Fit/Simulation_study/Output/fit_RV_SIRS_npi_vi_simul2_a_iav2.RDS")
```

```{r}
fit <- readRDS("Fit/Simulation_study/Output/fit_RV_SIRS_npi_vi_simul2_a_iav2.RDS")

check_hmc_diagnostics(fit)
get_num_divergent(fit)
get_num_max_treedepth(fit)
get_low_bfmi_chains(fit)
get_bfmi(fit)

summary(fit)$summary[c(1:5,53:60),]
```

```{r}
posterior <- extract(fit)

traceplot(fit, pars=parms)
pairs(fit, pars=parms)
plot_fit(posterior, data_stan)
plot_beta(posterior)
```